---
title: "R Primer: Markdown Example"
author: "Kara McCormack"
date: "2022-08-12"
output: html_document
---

```{r setup, echo = F, message = F, warning = F}
library(tidyverse)
library(kableExtra)
library(table1)
```

The following code shows some things we can do with R Markdown!

## Load the `mtcars` data

(Adapted from [https://www.datavis.ca/courses/RGraphics/R/gg-cars.html](https://www.datavis.ca/courses/RGraphics/R/gg-cars.html))

The `mtcars` data was extracted from the 1974 *Motor Trend* US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). 

```{r}
data(mtcars, package="datasets")
str(mtcars)
```
## Variable Definitions

```{r echo = F}
variable <- colnames(mtcars)
name <- c("Miles/(US) gallon", "Number of cylinders", "Displacement (cu.in.)", 
          "Gross horsepower", "Rear axle ratio", "Weight (1000 lbs)", 
          "1/4 mile time", "Engine (0=V-shaped, 1=straight)", 
          "Transmission (0 = auto, 1 = manual)", "Number of forward gears", 
          "Number of carburetors")
table <- data.frame(variable, name)

table %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


## Summary Statistics for select variables
```{r echo = F}
table1(~ mpg + disp + hp + drat + factor(cyl) +factor(vs) , data=mtcars) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

## Table options with kable(): select mtcars observations

```{r echo = F, message = F}
mtcars[1:8, 1:8] %>%
  kbl() %>%
  kable_paper(full_width = F) %>%
  column_spec(2, color = spec_color(mtcars$mpg[1:8]),
              link = "https://haozhu233.github.io/kableExtra/") %>%
  column_spec(6, color = "white",
              background = spec_color(mtcars$drat[1:8], end = 0.7),
              popover = paste("am:", mtcars$am[1:8]))
```

## Plots with ggplot2

```{r echo = F}
library(ggplot2)
ggplot(mtcars, aes(x=hp, y=mpg, color=as.factor(cyl), shape=as.factor(cyl))) +
    geom_point(size=3) +
  labs(xlab = "Horsepower", 
       ylab = "Miles/(US) gallon") +
  ggtitle("Miles/(US) gallon decreases with increased horsepower") +
  theme_bw()
```


## Resources

- [Markdown syntax](https://www.markdownguide.org/basic-syntax/)
- [Kable](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)
- [ggplot2](https://r-graph-gallery.com/ggplot2-package.html)

