---
title: "Exercises"
author: "Kara McCormack"
date: '2022-08-04'
output: html_document
---

```{r echo = F, warning = F, message = F}
# load libraries
library(babynames)
library(tidyverse)
```


## Exercise 1

Assign the numbers 3 through 12 to a vector called "a". Print a. 

```{r}
a <- 3:12
a
```

Run the code `a+3`. What happens?

```{r}
a+3
```

Find the length of the vector using the `length()` function, and find the mean of the vector using the `mean()` function.

```{r}

```

What will the code `a == 5` return? Think about it first, then run the code. 

```{r}

```


## Exercise 2

1. In your Console, print your working directory. 
2. Read in the iris_example.csv file using `read_csv(<FULL PATH TO FILE>)`, assigning it to a variable called `dat`. Then print the data.

```{r}

```



## Exercise 3

1. Filter the babynames dataset based on the following criteria: 
  - name is Jessica AND year is 2001
```{r}
babynames %>%
  filter(name == "Jessica" & year == 2001)
```

2. Filter the babynames dataset to the following names: Anna, Alice, or Sarah. 
```{r}
babynames %>%
  filter(name %in% c("Anna", "Alice", "Sarah"))
```

3. Filter the mpg dataset to cars getting at least 19 city miles per gallon and at least 30 highway miles per gallon.
```{r}
mpg %>%
  filter(cty >= 19 & hwy >= 30)
```


## Grouping

```{r}
pollution <- tribble(
       ~city,   ~size, ~amount,
  "New York", "large",      23, 
  "New York", "small",      14, 
  "London",   "large",      22,
  "London",   "small",      16, 
  "Beijing",  "large",     121, 
  "Beijing",  "small",     56
)
```

## Exercise 4

Group the babynames dataset by sex and then find out the total number of names for each group. 

```{r}

```


## Exercise 5

Group the babynames dataset by name and sex, find the total number of names for each combination, and then arrange the totals from largest to smallest. 

```{r}

```

## Exercise 6

Plot the code on the previous slide (creating a barplot showing the top ten most common names), but replacing `x = fct_reorder(name, desc(total))` with `x = name`. What happened? Why is `fct_reorder()` helpful?


## Exercise 6

In your groups:

What relationship do you expect to see between engine size (displ) and mileage (hwy)?

## Exercise 7

- Among your groups, run the following code in a new chunk. 

- What does each line within the `labs()` function do? 

```{r}
mpg %>%
  ggplot(mapping = aes(displ, hwy, color = class)) +
  geom_point() +
  labs(title = "Fuel Efficiency by Engine Size",
       subtitle = "Data facetted by class",
       x = "Engine Size (displacement in liters)",
       y = "Fuel Efficiency (MPG)",
       color = "Class of\nAutomobile",
       caption = "Data from the EPA")
```

## Exercise 8

```{r}

```

Change the code for the above plot to produce boxplots instead. Bonus: add an informative title. 
```{r}

```

## Exercise 9

Make a bar chart of `class` below. Use the cheat sheet and/or google. Hint: do not supply a `y` variable. 

```{r}

```

## Modeling

```{r}
library(modelr)
```

```{r}
wages <- heights %>%
  filter(income > 0)
```
## Exercise 10

Fit the following model: `mod_e <- lm(log(income) ~ education, data = wages)`
Examine the output. What does it look like? Use the `summary` function to look at the full output of the model. 
```{r}

```

## Exercise 11

Use a pipe to model `log(income)` against `height`. Then use broom and dplyr to extract:

1. The coefficient estimates and their related statistics. 

2. The `adj.r.squared` and `p.value` for the overall model.

```{r}

```

## Exercise 12

Model `log(income)` against `education` and `height` and `sex`. Can you interpret the coefficients?

```{r}

```

