---
title: "05_class_code"
author: "Kara McCormack"
date: "2022-08-16"
output: html_document
---

```{r setup}
library(tidyverse)
library(babynames)
library(plotly)
library(modelr)
library(broom)
```


### Slide 41: Some math

```{r}
5 + 7
3.14*3
```

### Slide 43: R is a functional language
```{r}
3
"California"
"2015-03-07"

x <- c(5/6, 0.83, 7)
round(x, digits = 3)
```

### Slide 45: Example: Scalars
```{r}
s1 <- 2       # numeric
s2 <- "Hi!"   # character
s3 <- TRUE    # logical

s1
s2
s3
```

### Slide 46: Vectors (stacks of scalars)

```{r}
a <- rep(1,4)
b <- seq(1, 2, by = .2)
a
b
c <- c(a, b)
d <- 1:4
d
```

### Slide 47: Matrices (stacks of vectors)

```{r}
M1 <- matrix(1:6, 3, 2)
M2 <- matrix(1:6, 3, 2, byrow = TRUE)

M1
M2
```

### Slide 49: Example: data frames

```{r}
num <- 1:4
chr <- c("M", "F", "F", "M")
log <- c(TRUE, TRUE, FALSE, TRUE)
D1 <- data.frame(num, chr, log)
D1
```

### Slide 50: Example: vectorization

```{r}
a <- c(1, 2, 3)
b <- c(3, 4, 5)
a + b
a * b
```

### Slide 51: Example: vectorization

```{r}
a <- c(1, 2, 3)
b <- c(3, 4)
a+b  # warning message since a and b are different lengths
```

### Slide 52: Example: functions

```{r}
values <- c(1:10, 17)
print(values)
mean(values)
median(values)
sd(values)
```


### Slide 54: Example

```{r}
c <- 1:5
d <- c(0, 1, 7, 7, 7)
c < d
c == d
```

### Slide 72: Example

```{r}
# example_data <- read_csv("04_data.csv")
# example_data

# change to your working directory if not in project folder
```

### Slide 77: select(): example

```{r}
select(babynames, name, prop)
```

### Slide 78: select(): example with pipe

```{r}
babynames %>%
  select(name, prop)
```

### Slide 79: select(): helper functions

```{r}
colnames(mpg)
mpg %>%
  select(cty:class)
```

### Slide 80: select(): helper functions

```{r}
mpg %>%
  select(-c(cty, hwy))
```

### Slide 81: select(): helper functions

```{r}
mpg %>%
  select(starts_with("c"))
```

### Slide 82: select(): helper functions

```{r}
mpg %>%
  select(ends_with("y"))
```

### Slide 84: filter(): example

```{r}
babynames %>%
  filter(name == "Garrett")
```

### Slide 86: Two common mistakes

```{r}
# using = instead of ==
# babynames %>%
#   filter(name = "Sea")      # incorrect

babynames %>%
  filter(name == "Sea")     # correct

# forgetting quotes
# babynames %>%
#   filter(name = Sea)      # incorrect

babynames %>%
  filter(name == "Sea")     # correct
```

### Slide 87: filter()

```{r}
babynames %>%
  filter(name == "Garrett", year == 1880)
```

### Slide 88: filter() for multiple conditions

```{r}
babynames %>%
  filter(name == "Garrett" & year == 1880)
```

### Slide 90: Boolean operators: example

```{r}
x <- 2.5
x >=2
x < 3
x >= 2 & x < 3
x != 5
```

### Slide 91: filter() with conditions

```{r}
babynames %>%
  filter(name == "Sue", sex == "M")
```

### Slide 92: filter() with conditions

```{r}
babynames %>%
  filter((n == 5 | n == 6) & year == 1880)
```

### Slide 93: More common mistakes

```{r}
# babynames %>%
#   filter(10 < n < 20)    # error

babynames %>%
  filter(10 < n, n < 20) # correct
```

### Slide 94: More common mistakes

```{r}
# will run, just not efficient
babynames %>%
  filter(n == 5 | n ==6 | n == 7 | n == 8) 

# better
babynames %>%
  filter(n %in% c(5, 6, 7, 8))

# or, alternatively:
vec <- 5:8
babynames %>%
  filter(n %in% vec)
```

### Slide 96: arrange()

```{r}
babynames %>%
  arrange(n)
```

### Slide 97: arrange()

```{r}
babynames %>%
  arrange(desc(n))
```

### Slide 98: Performing multiple steps

```{r}
boys_2015 <- babynames %>% filter(year == 2015, sex == "M")
boys_2015 <- boys_2015 %>% select(name, n)
boys_2015 <- boys_2015 %>% arrange(desc(n))
boys_2015
```

### Slide 99: Performing multiple steps: nesting functions

```{r}
# not recommended
arrange(select(filter(babynames, year == 2015, sex == "M"), name, n), desc(n))
```

### Slide 100: Performing multiple steps: piping each step

```{r}
babynames %>%
  filter(year == 2015, sex == "M") %>%
  select(name, n) %>%
  arrange(desc(n))
```

### Slide 101: What if we wanted to visualize

```{r}
babynames %>%
  filter(name == "Garrett", sex == "M") %>%
  select(year, prop) %>%
  ggplot() +
  geom_line(mapping = aes(year, prop))
```


### Slide 103: Plotting groups

```{r}
babynames %>%
  filter(name == "Michael") %>%
  ggplot() +
  geom_point(mapping = aes(year, prop))
```

### Slide 104: Plotting groups, adding in color

```{r}
babynames %>%
  filter(name == "Michael") %>%
  ggplot() +
  geom_line(mapping = aes(year, prop, color = sex))
```

### Slide 105: Plotting with facet_wrap()

```{r}
babynames %>%
  filter(name == "Michael") %>%
  ggplot() +
  geom_line(mapping = aes(year, prop)) +
  facet_wrap(~sex)
```

### Slide 107: summarize()

```{r}
babynames %>%
  head(3)

babynames %>%
  summarize(total = sum(n), max = max(n))
```

### Slide 108: n()

```{r}
babynames %>%
  head(3)

babynames %>%
  summarize(n = n())
```

### Slide 109: n_distinct()

```{r}
babynames %>%
  head(3)

babynames %>%
  summarize(n = n(), nname = n_distinct(name))
```

### Slide 111: Name popularity example: Khaleesi

```{r}
babynames %>%
  filter(name == "Khaleesi" & sex == "F")
```

### Slide 112: Name popularity example: Khaleesi

```{r}
babynames %>%
  filter(name == "Khaleesi" & sex == "F") %>%
  summarize(total = sum(n))
```

### Slide 114: Grouping cases

```{r}
pollution <- tribble(
       ~city,   ~size, ~amount,
  "New York", "large",      23, 
  "New York", "small",      14, 
  "London",   "large",      22,
  "London",   "small",      16, 
  "Beijing",  "large",     121, 
  "Beijing",  "small",     56
)
```

### Slide 115: grouping

```{r}
pollution %>%
  summarize(mean = mean(amount), 
            sum = sum(amount), 
            n = n())
```

### Slide 116: group_by()

```{r}
pollution %>%
  group_by(city)
```

### Slide 117: group_by()

```{r}
pollution %>%
  group_by(city) %>%
  summarize(mean = mean(amount), 
            sum = sum(amount), 
            n = n())
```

### Slide 118: group_by()

```{r}
pollution %>%
  group_by(city, size) %>%
  summarize(mean = mean(amount), 
            sum = sum(amount), 
            n = n())
```

### Slide 121: ungroup()

```{r}
babynames %>%
  group_by(sex) %>%
  ungroup() %>%
  summarize(total = sum(n))
```

### Slide 125: Visualizing the number of children over time

```{r}
babynames %>%
  group_by(year) %>%
  summarize(n_children = sum(n)) %>%
  ggplot() +
  geom_line(mapping = aes(x= year, y = n_children))
```

### Slide 127: mutate()

```{r}
babynames %>%
  mutate(percent = round(prop*100, 2))
```

### Slide 128: mutate()

```{r}
babynames %>%
  mutate(percent = round(prop*100, 2), 
         nper = round(percent))
```

### Slide 133: Data exploration: mpg

```{r}
mpg
```

### Slide 138: Visualize data with ggplot2

```{r}
ggplot(mpg) + geom_point(aes(x = displ, y = hwy, color = class))
ggplot(mpg) + geom_point(aes(x = displ, y = hwy, size = class))
ggplot(mpg) + geom_point(aes(x = displ, y = hwy, shape = class))
ggplot(mpg) + geom_point(aes(x = displ, y = hwy, alpha = class))
```

### Slide 140: Visualize data with ggplot2

```{r}
mpg %>%
  ggplot(mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point()
```

### Slide 142: facets

```{r}
mpg %>%
  ggplot(mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  facet_wrap(~class)
```

### Slide 154: filling with a new variable

```{r}
mpg %>%
  ggplot() +
  geom_bar(aes(x = class, fill = drv))
```


### Slide 155-156: stacking layers

```{r}
mpg %>%
  ggplot() +
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

### Slide 158: changing colors for different geom layers
```{r}
mpg %>%
  ggplot(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth()
```

### Slide 159: global vs. local 
```{r}
mpg %>%
  ggplot(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(data = filter(mpg, drv == "r"))
```

### Slide 162-163: interactivity
```{r}
p <- ggplot(mpg) +
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))

# library(plotly)
ggplotly(p)
```

### Slide 176: Modeling example dataset from modelr package
```{r}
wages <- heights %>%
  filter(income > 0 )
```

### Slide 177: wages dataset
```{r}
wages %>%
  slice(1:6) %>%
  knitr::kable()
```

### Slide 178: wages dataset (continued)
```{r}
wages %>%
  ggplot(aes(income)) +
  geom_histogram()
```

### Slide 179: wages data: log transformation
```{r}
wages %>%
  ggplot(aes(income)) +
  geom_histogram(binwidth = 0.25)
```

### Slide 181: lm()
```{r}
lm(log(income) ~ education, data = wages)
```

### Slide 184: .
```{r}
mod_e <- wages %>%
  lm(log(income) ~ education, data = .)
```

### Slide 187: tidy()
```{r}
mod_e <- wages %>%
  lm(log(income) ~ education, data = .)
mod_e %>%
  tidy()
```

### Slide 188: glance()
```{r}
mod_e %>%
  glance()
```

### Slide 189: augment()
```{r}
mod_e %>%
  augment()
```

### Slide 192: comparing models
```{r}
# running model from previous exercise
mod_h <- wages %>%
  lm(log(income) ~ height, data = .)

mod_h %>%
  tidy() %>%
  filter(p.value < 0.05)

mod_e %>%
  tidy() %>%
  filter(p.value < 0.05)
```

### Slide 194: multivariate regression
```{r}
log(income) ~ education + height
```

### Slide 195: modeling education and height
```{r}
mod_eh <- wages %>%
  lm(log(income) ~ education + height, data = .)
mod_eh %>%
  tidy()
```


### Slide 200: model visualization
```{r}
wages %>%
  ggplot(aes(x = height, y = log(income))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm)      # fits y ~ x (mod_h)
```

### Slide 203: add_predictions() for height lm
```{r}
wages %>%
  add_predictions(mod_h)
```


### Slide 205: add_residuals()
```{r}
wages %>%
  add_residuals(mod_h)
```


### Slide 209: spread_residuals()
```{r}
# mod_ehs from Exercise 13
mod_ehs <- wages %>%
  lm(log(income) ~ education + height + sex, data = .)
wages %>%
  spread_residuals(mod_h, mod_eh, mod_ehs)
```

### Slide 211: gather_residuals()
```{r}
# mod_ehs from Exercise 13
wages %>%
  gather_residuals(mod_h, mod_eh, mod_ehs) %>%
  arrange(income, height)
```