---
title: "Exercises_Solutions"
author: "Kara McCormack"
date: '2022-08-08'
output: html_document
---

## Exercise 1

Assign the numbers 3 through 12 to a vector called "a". Print a. 

```{r}
a <- 3:12
a
```

Run the code `a+3`. What happens?

```{r}
a+3
```

Find the length of the vector using the `length()` function, and find the mean of the vector using the `mean()` function.

```{r}

```

What will the code `a == 5` return? Think about it first, then run the code. 

```{r}

```

## Exercise 2

Read in the iris_example.csv file using `read_csv`, assigning it to a variable called `dat`. Then print the data. 
```{r}
dat <- read_csv("~/school/rprimer/iris_example.csv")
dat
```

What is the dimension of "dat"? (Hint: use the `dim()` function). 
```{r}
dim(dat)
```

What are the column names of "dat"? (Hint: use the `colnames()` function). 
```{r}
colnames(dat)
```


## Exercise 3

1. Filter the babynames dataset based on the following criteria: 
  - name is Jessica AND year is 2001
```{r}
babynames %>%
  filter(name == "Jessica" & year == 2001)
```

2. Filter the babynames dataset to the following names: Anna, Alice, or Sarah. 
```{r}
babynames %>%
  filter(name %in% c("Anna", "Alice", "Sarah"))
```



## Exercise 4

Group the babynames dataset by sex and then find out the total number of names for each group. 

```{r}
babynames %>%
  group_by(sex) %>%
  summarize(total = sum(n))
```


## Exercise 5

Group the babynames dataset by name and sex, find the total number of names for each combination, and then arrange the totals from largest to smallest. 

```{r}
babynames %>%
  group_by(name, sex) %>% 
  summarise(total = sum(n)) %>% 
  arrange(desc(total))
```

## Exercise 6

Go through each of the lines of the code on the previous slide, noting what each line does. 
Plot the code on the previous slide (creating a barplot showing the top ten most common names), but replacing `x = fct_reorder(name, desc(total))` with `x = name`. What happened? Why is `fct_reorder()` helpful?

```{r}
babynames %>%
  group_by(name, sex) %>% 
  summarise(total = sum(n)) %>% 
  arrange(desc(total)) %>% 
  ungroup() %>% 
  slice(1:10) %>% 
  ggplot() +
    geom_col(mapping = aes(x = name, 
             y = total, fill = sex)) +
    theme_bw() +
    scale_fill_brewer() +
    labs(x = "name")
```

## Exercise 6

In your groups:

What relationship do you expect to see between engine size (displ) and mileage (hwy)?

We might expect that larger engines can get better mileage. 

## Exercise 7

The labs() function allows us to print a title, subtitle (etc) , and name for the legend. Use `\n` for a new line. 

## Exercise 8

```{r}
mpg %>%
  ggplot() +
  aes(x = class, y = hwy) +
  geom_boxplot() +
  ggtitle("Boxplots of class vs mileage (hwy) for US cars")
```

## Exercise 9

Make a bar chart of `class` below. Use the cheat sheet and/or google. Hint: do not supply a `y` variable. 

```{r}
mpg %>%
  ggplot() +
  geom_bar(aes(x = class, fill = class))
```
