---
title: "R Primer: Markdown Example"
author: "Kara McCormack"
date: "2022-08-12"
output: html_document
---

```{r echo = F, message = F, warning = F}
library(tidyverse)
library(kableExtra)
library(table1)
```

The following code shows some things we can do with R Markdown!

## Load the `mtcars` data

(Adapted from [https://www.datavis.ca/courses/RGraphics/R/gg-cars.html](https://www.datavis.ca/courses/RGraphics/R/gg-cars.html))

The `mtcars` data was extracted from the 1974 *Motor Trend* US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). 

```{r}
data(mtcars, package="datasets")
str(mtcars)
```
## Variable Definitions

```{r echo = F}
variable <- colnames(mtcars)
name <- c("Miles/(US) gallon", "Number of cylinders", "Displacement (cu.in.)", 
          "Gross horsepower", "Rear axle ratio", "Weight (1000 lbs)", 
          "1/4 mile time", "Engine (0=V-shaped, 1=straight)", 
          "Transmission (0 = auto, 1 = manual)", "Number of forward gears", 
          "Number of carburetors")
table <- data.frame(variable, name)

table %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


## Summary Statistics for select variables
```{r echo = F}
table1(~ mpg + disp + hp + drat + factor(cyl) +factor(vs) , data=mtcars) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

## Table options with kable(): select mtcars observations

```{r echo = F, message = F}
mtcars[1:8, 1:8] %>%
  kbl() %>%
  kable_paper(full_width = F) %>%
  column_spec(2, color = spec_color(mtcars$mpg[1:8]),
              link = "https://haozhu233.github.io/kableExtra/") %>%
  column_spec(6, color = "white",
              background = spec_color(mtcars$drat[1:8], end = 0.7),
              popover = paste("am:", mtcars$am[1:8]))
```

## Plots with ggplot2

```{r echo = F}
library(ggplot2)
ggplot(mtcars, aes(x=hp, y=mpg, color=as.factor(cyl), shape=as.factor(cyl))) +
    geom_point(size=3) +
  labs(xlab = "Horsepower", 
       ylab = "Miles/(US) gallon") +
  ggtitle("Miles/(US) gallon decreases with increased horsepower") +
  theme_bw()
```

## Mapping with ggplot2()

```{r echo = F, message = F}

# Geospatial data available at the geojson format
library(geojsonio)
spdf <- geojson_read("https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",  what = "sp")

# Since it is a bit too much data, I select only a subset of it:
spdf <- spdf[ substr(spdf@data$code,1,2)  %in% c("06", "83", "13", "30", "34", "11", "66") , ]

# read data
data <- read.table("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/data_on_french_states.csv", header=T, sep=";")

# I need to fortify the data AND keep trace of the commune code! (Takes ~2 minutes)
library(broom)
spdf_fortified <- tidy(spdf, region = "code")

# Make the merge
spdf_fortified = spdf_fortified %>%
  left_join(. , data, by=c("id"="depcom"))

# Note that if the number of restaurant is NA, it is in fact 0
spdf_fortified$nb_equip[ is.na(spdf_fortified$nb_equip)] = 0.001

library(viridis)
p <- ggplot() +
  geom_polygon(data = spdf_fortified, aes(fill = nb_equip, x = long, y = lat, group = group) , size=0, alpha=0.9) +
  theme_void() +
  scale_fill_viridis(trans = "log", breaks=c(1,5,10,20,50,100), name="Number of restaurant", guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) ) +
  labs(
    title = "South of France Restaurant concentration",
    subtitle = "Number of restaurant per city district",
    caption = "Data: INSEE | Creation: Yan Holtz | r-graph-gallery.com"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),

    plot.title = element_text(size= 22, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    plot.subtitle = element_text(size= 17, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.43, l = 2, unit = "cm")),
    plot.caption = element_text( size=12, color = "#4e4d47", margin = margin(b = 0.3, r=-99, unit = "cm") ),

    legend.position = c(0.7, 0.09)
  ) +
  coord_map()
p
```



```{r echo = F, message = F, warning = F}
# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```


### Watercolor theme 

```{r echo = F, message = F}
# Library
library(ggmap)
library(gridExtra)
 
# For stamen map, you have to give the extremity of the window you are looking at. here is an example with the watercolor background (Around brisbane)
map <- get_stamenmap( bbox = c(left = 110, bottom = -40, right = 160, top = -10), zoom = 4, maptype = "watercolor")
ggmap(map) + 
  theme_void() + 
  theme(
    plot.title = element_text(colour = "orange"), 
    panel.border = element_rect(colour = "grey", fill=NA, size=2)
  )
```

## Resources

- [Markdown syntax](https://www.markdownguide.org/basic-syntax/)
- [Kable](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)
- [ggplot2](https://r-graph-gallery.com/ggplot2-package.html)

